---
title: "Superstore Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: ["twitter", "facebook", "menu"]
    theme: readable
---


```{r setup, include=FALSE, warning=TRUE, message=FALSE}
# R packages indstalled 

library(flexdashboard)
library(tidyverse)
library(lubridate)
library(skimr)
library(fpp2)
library(knitr)
library(DT)
library(rpivotTable)
library(plotly)
library(openintro)
library(highcharter)


# Data import
ss <- read.csv("superstore2.csv", header = T)

# Data manipulation
ss <- ss %>% 
  mutate(Order.Date = ymd(Order.Date),
         Ship.Date = ymd(Ship.Date),
         Year = factor(Year),
         Month = factor(Month),
         Ship.Mode = factor(Ship.Mode),
         Customer.Name =! factor(Customer.Name),
         Segment = factor(Segment),
         Country = factor(Country),
         City = factor(City),
         State = factor(State),
         Region = factor(Region),
         Category = factor(Category),
         Sub.Category = factor(Sub.Category),
         Product.Name = factor(Product.Name)
         ) %>%    # These are datas and factor conversion          
  select(-X, -Ship.Date, -Product.ID, -Postal.Code, -Country, -Region, -Year, -Month, -date.var) %>%    # Removing some unnecessary variables to simplify the df
  select(State, City, Order.Date, Ship.Mode, Customer.Name, Segment, Category, Sub.Category, Product.Name, Sales) # Rearranging all variables
 
# Preparing some additional tables  
# Furniture df
furni <- ss %>% filter(Category == "Furniture")

# Office suppliesdf
office <- ss %>% filter(Category == "Office Supplies")

# Technology df
tech <- ss %>% filter(Category == "Technology")

```


Executive Summary {data-icon="fa-users"}
=====================================================================================

R1
----------------------------------------------------------

### Last Update

```{r}
valueBox(Sys.time(),
         icon = "glyphicon-time", color = "lightgrey")
```



### Total Sales

```{r}
valueBox(paste0("$", prettyNum(sum(ss$Sales), big.mark = ",")),
         icon = "fa-dollar-sign")
 

```

### Total Orders

```{r}
valueBox(paste0(prettyNum(nrow(ss), ",")),
         icon = "fa-file-invoice")


```

### Average Sales per Order

```{r}
a <- ss %>% 
  summarise(count = n(),
                 total.sales = sum(Sales)) 

ave.sales <- round(a$total.sales/a$count)

valueBox(paste0("$", ave.sales), 
         icon = "fab fa-circle-dollar")

```



R1
----------------------

### Product Type

```{r, message=FALSE, warning=FALSE}

library(networkD3)

networkData <- ss %>% 
  select(Category, Sub.Category) %>% 
  group_by(Category, Sub.Category) %>% 
  summarise(count = n()) 

simpleNetwork(networkData, zoom = T, 
              linkDistance = 80,
              fontSize = 15,
              opacity = 0.8,
              width = 500
              )

```

### Product Sales

```{r, warning=FALSE, message=FALSE}
# set df
bubble.data <- ss %>% 
  select(Category, Sub.Category, Sales) %>% 
  group_by(Category, Sub.Category) %>% 
  summarise(total.sales = round(sum(Sales)),
            total.sales2 = prettyNum(total.sales, big.mark = ","))

# chart
hc <- hchart(object = bubble.data, "packedbubble",
       hcaes(name = "Sub.Category", value = "total.sales", group = "Category"))

q60 <- as.numeric(quantile(bubble.data$total.sales, .60))   # Only catgeory >0.95 will be appear

hc %>% 
  hc_tooltip(
    useHTML = TRUE,
    pointFormat = "<b>{point.name}:</b> {point.value}"
  ) %>% 
  hc_plotOptions(
    packedbubble = list(
      maxSize = "50%",
      zMin = 0,
      layoutAlgorithm = list(
        gravitationalConstant =  0.05,
        splitSeries =  TRUE, # TRUE to group points
        seriesInteraction = TRUE,
        dragBetweenSeries = TRUE,
        parentNodeLimit = TRUE
      ),
      dataLabels = list(
        enabled = TRUE,
        format = "{point.name}",
        filter = list(
          property = "y",
          operator = ">",
          value = q60
        ),
        style = list(
          color = "black",
          textOutline = "none",
          fontWeight = "normal"
        )
      )
    )
  )
```


R1  
-----------------------------

### Best Sellers

```{r}
library(collapsibleTree)

tree.data <- ss %>% 
  group_by(Segment, Category, Sub.Category) %>% 
  summarise(total.Sales = round(sum(Sales)))

collapsibleTreeSummary(tree.data,
                root = "Superstore",  
                hierarchy = c("Segment", "Category", "Sub.Category"),  
                width = 800, 
                height = 1000,
                attribute = "total.Sales",
                zoomable = T,
                collapsed = F, 
                fontSize = 15, 
                nodeSize = "total.Sales")  


```


Furniture {data-navmenu="Menu A"}
===============================

R2 
-------------------

### Sales Analysis (2015 - 2018)

```{r}
valueBox(paste("Furniture"),
         color = "orange")

```

### Total Sales 

```{r}
valueBox(paste0("$", prettyNum(sum(furni$Sales), big.mark = ",")),
         icon = "fa-money")

```

### Proportion of Grand Total Sales 

```{r}

gauge(value = sum(furni$Sales),  
      min = 0,  
      max = sum(ss$Sales))  

```



### 2015 Total Sales 

```{r}
valueBox(paste0("$", prettyNum(156478, big.mark = ",")),
         icon = "fa-money",
         color = "grey")

```

### 2016 Total Sales 

```{r}
valueBox(paste0("$", prettyNum(164054	, big.mark = ",")),
         icon = "fa-money",
         color = "grey")

```

### 2017 Total Sales 

```{r}
valueBox(paste0("$", prettyNum(195813, big.mark = ",")),
         icon = "fa-money",
         color = "grey")

```

### 2018 Total Sales 

```{r}
valueBox(paste0("$", prettyNum(212313, big.mark = ",")),
         icon = "fa-money",
         color = "grey")

```

R2 {data-height = 350}
----------------------------------------

### Daily Sales

```{r}
g1 <- ggplot(furni, aes(x = Order.Date, y = Sales)) + geom_point() + geom_line() +
  labs(x = "Date", y = "Sales, $") + theme_minimal()

ggplotly(g1)

```

### Daily Sales (Log)

```{r}
g2 <- ggplot(furni, aes(x = Order.Date, y = log(Sales))) + geom_point() + geom_line() +
  labs(x = "Date", y = "Sales, $") + theme_minimal()

ggplotly(g2)

```

### Monthly Sales

```{r}
# Set up df

furni2 <- furni %>% 
  mutate(Date = paste0(year(Order.Date), "-", month(Order.Date)),
         Date = ym(Date)) %>% 
  group_by(Date) %>% 
  summarise(montly.sales = sum(Sales))

furni.ts <- ts(furni2$montly.sales, start = c(2015,1), frequency = 12)

# qplot

g3 <- autoplot(furni.ts) + 
  labs(y = "Sales, $") +
  geom_point() +
  geom_vline(aes(xintercept = 2016), color = "purple", linetype = 2) +
  geom_vline(aes(xintercept = 2017), color = "purple", linetype = 2) +
  geom_vline(aes(xintercept = 2018), color = "purple", linetype = 2) +
  geom_vline(aes(xintercept = 2019), color = "purple", linetype = 2) +
  geom_vline(aes(xintercept = 2015), color = "purple", linetype = 2) +
  theme_classic()

ggplotly(g3)

```


R2 {data-height = 650}
----------------

### Difference Plot

```{r}
furni.ts <- ts(furni2$montly.sales, start = c(2015,1), frequency = 12)
furni.ts.d <- diff(furni.ts)

g4 <- autoplot(furni.ts.d) + geom_point(color = "orange3") + geom_line(color = "orange") + theme_classic() +
  labs(y = "Sales, $") 

ggplotly(g4)

```

### Seasonal Plot

```{r}
g5 <- ggseasonplot(furni.ts.d) + geom_point() + theme_classic() + labs(y = "Sales, $") 
ggplotly(g5)
```

### Subseries Plot

```{r}
g6 <- ggsubseriesplot(furni.ts.d) + geom_point(color = "orange2") + geom_line(color = "orange2") +
  theme_classic() + labs(y = "Sales, $") 
ggplotly(g6)

```


Furniture Forecast {data-orientation=columns}{data-navmenu="Menu A"}
==========================================================================


### ETS Forecast of Next 24 Months

```{r}
library(highcharter)

furni_ets <- ets(furni.ts)
furni_ets_fc <- forecast(furni_ets, h = 24, level = 95)

hchart(furni_ets_fc)


```

### Model Summary

```{r}
summary(furni_ets_fc)

```


### Model Diagnostic 

```{r}
checkresiduals(furni_ets_fc)


```


Map
=======================================

R3
-------------------------

### Overall Sales

```{r}
overall <- ss %>% group_by(State) %>% summarise(total.sales = sum(Sales))
  
highchart() %>% 
  hc_add_series_map(usgeojson, overall,
                    name = "State",
                    value = "total.sales",
                    joinBy = c("woename", "State")) %>%
  hc_mapNavigation(enabled = T) %>% 
  hc_colorAxis(minColor = "white", maxColor = "Black")


```
R3
-------------------------

### Furniture sales

```{r}
f <- furni %>% group_by(State) %>% summarise(total.sales = sum(Sales))

highchart() %>% 
  hc_add_series_map(usgeojson, f,
                    name = "State",
                    value = "total.sales",
                    joinBy = c("woename", "State")) %>%   # woe = where on earth
  hc_mapNavigation(enabled = T) 
  
```
### Office Supplies sales 

```{r}

o <- office %>% group_by(State) %>% summarise(total.sales = sum(Sales))

highchart() %>% 
  hc_add_series_map(usgeojson, o,
                    name = "State",
                    value = "total.sales",
                    joinBy = c("woename", "State")) %>%   # woe = where on earth
  hc_mapNavigation(enabled = T) %>% 
  hc_colorAxis(minColor = "yellow", maxColor = "green")
  
```

### Technology sales 

```{r}
t <- office %>% group_by(State) %>% summarise(total.sales = sum(Sales))

highchart() %>% 
  hc_add_series_map(usgeojson, t,
                    name = "State",
                    value = "total.sales",
                    joinBy = c("woename", "State")) %>%   # woe = where on earth
  hc_mapNavigation(enabled = T) %>% 
  hc_colorAxis(minColor = "orange", maxColor = "brown")
  
```

Raw Data
===============================


```{r}
datatable(ss,
          caption = "Raw Sales Data",
          rownames = T,
          filter = "top",
          option = list(pageLength = 25))

```

Pivot Table
=================================================================================

```{r}
# Data frame

furni2 <- furni %>% 
  mutate(Date = paste0(year(Order.Date), "-", month(Order.Date)),
         Date = ym(Date)) %>% 
  group_by(Date,Ship.Mode, Segment, State, Category, Sub.Category) %>% 
  summarise(montly.sales = sum(Sales))

# Interactive Pivot table 

rpivotTable(furni2,
            rendererName = "Heatmap",
            cols = "Order.Date",
            rows = "State")
```


